---
title: "Course Project - Logistic regression"
author: "Jaeyoung Lee"
date: "11/29/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(data.table)
library(tidyverse)
library(bit64)
library(lubridate)
```


```{r load_data, eval=FALSE}
# Load data
tweetdata <- fread('uselection_tweets_1jul_11nov.csv')

# Select variables of interest
# Remove redundant varaiables (Negativity, Positivity) b/c there is score
tweetdata_select <- tweetdata %>% select(`Retweet-Count`, PartyName, Score)
names(tweetdata_select) <- c('retweet_count', 'party', 'score')

saveRDS(tweetdata_select, 'tweetdata_select.RDS')

```

```{r tidy_data, eval=FALSE}
# Load RDS data
twitter_data <- readRDS('tweetdata_select.RDS')

# Remove observations with zero retweet_count 
# Left observations mentioned one of two parties only: Republicans or Democrats
twitter <- twitter_data %>% filter(retweet_count != 0) %>% 
  filter(party == 'Republicans' | party == 'Democrats') 

```

```{r logistic regression, eval = FALSE}
##### Multinomial Logistic regression with bootstrap-like method #####
# Sample from the data
set.seed(20202021)
n <- 1000 # Sample size
M <- 1000   # Number of iteration

logit_coefficients <- NULL # Coefficients from M logit models
# Generate M coefficients
for(i in 1:M){
  # Sample from data with size n
  index <- sample(1:nrow(twitter), size = n) 
  twitter_sample <- twitter[index,]           
  
  # Response variable : Democrats, Republicans
  logit_fit <- nnet::multinom(party~retweet_count + score, 
                              data = twitter_sample, trace = FALSE)
  logit_coefficients <- rbind(logit_coefficients, summary(logit_fit)$coefficients)

}

# Save RData
save.image('logit_model.RData')


```

```{r load_rdata}
# Load RData
load('logit_model.RData')

```

```{r result}
# Summary of data used
print('Summary of data')
summary(twitter)

# Summary of logistic regression coefficients
print('Summary of logistic regression coefficients')
summary(logit_coefficients)

# Exponentiate regression coefficients
print('Summary of exponenetiated coefficients')
summary(exp(logit_coefficients))

```
\pagebreak

The raw data has $24201654$ observations. To reduce the size of data and to fix the right tail distribution of *Re-tweet counts*, the observations which have zero counts are excluded. Also, to control the multicollinearity, redundant variables are not selected for constructing a model. 

From the raw data, $12515164$ observations are used which *Re-tweet counts* are non-zero. Since the observations are still too large, it is necessary to use some methods handling the large data. To speed up computation and deal with large data, a re-sampling method is used which is similar to Bootstrap. Unlike Bootstrap, we sample without replacement and the sample size of each sample is smaller than the original.

The algorithm for the re-sampling method is as follows:

\begin{enumerate}
  \item For j = 1, \ldots, $M = 1000$,
    \begin{enumerate}
      \item Obtain sample of size $n = 1000$.
      \item Fit logistic regression from the $j$th sample.
      \item Store the coefficients of the model.
    \end{enumerate}
\end{enumerate}

From the $M = 1000$ samples, we can obtain a matrix of logistic regression coefficients. 

From the result, we can build a logistic regression model as follows:

\begin{eqnarray} 
\log \frac{P(Y = Republicans)}{P(Y = Democrats)} = 1.0995 - 4 \times 10^{-6} Retweet -0.27241 Score
\end{eqnarray}
where the baseline is *Democrats*, and the odds is defined as the proportion of Republicans to Democrats.

From the model, we can notice that the input *Re-tweet counts* is not significant. In other words, there is quite little effect of *Re-tweet counts* to the model. In addition, under fixed *Re-tweet counts*, the odds of Republicans is $exp(1.0995 -0.27241 \times Score) = 3.012 \times exp(-0.27241 \times Score)$. Therefore, as *Score* increases as one, the odds of Republicans decreases 23.84%. In other words, we can say that increase of the sentiment score of the tweets leads to the decrease of the odds of Republicans, and reversely, the increase of the odds of Democrats.





\pagebreak

```{r appendix, ref.label=knitr::all_labels(), echo = TRUE, eval=FALSE, include=TRUE}
